### 2007-01-03 Documentation/part1.tex
Part of manual updated

@@

\begin{table}[tpb]
\begin{center}
\begin{tabular}{p{3.8cm}p{3.8cm}p{7cm}}
\hline
Function & Purpose & Parameters \ra\ Result \\
\hline
\multicolumn{3}{l}{{\bf Systematics only:}} \\
\GLB{ChiSys} & $\chi^2$ with systematics only  & ({\tt glb\_params in, int exp, int rule}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projections onto axes:}} \\
\GLB{ChiTheta} & Projection onto $\theta_{13}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiDelta} & Projection onto $\deltacp$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiTheta23} & Projection onto $\theta_{23}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiDm} & Projection onto $\ldm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiDms} & Projection onto $\sdm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projection onto plane:}} \\
\GLB{ChiThetaDelta} & Projection onto $\theta_{13}$-$\deltacp$-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projection onto any hyper-plane:}} \\
\GLB{ChiNP} & Projection onto any $n$-dimensional hyper-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \newline
Needs \GLB{SetProjection} before! \\[0.2cm]
\multicolumn{3}{l}{{\bf Localization of degeneracies:}} \\
\GLB{ChiAll} & (Local) Minimization over all parameters  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\
\hline
\end{tabular}
\end{center}
\caption{\label{tab:stdfunctions} \index{norm}{Standard functions (table)} The \GLOBES\ standard function to obtain a $\chi^2$-value with systematics only or systematics and correlations. The parameters {\tt rule} and {\tt exp}
can either be \GLBC{GLB\_ALL} for all initialized experiment or the
experiment number ($0$ to \GLB{\_num\_of\_exps}-1) for a specific experiment. The format of \GLB{\_params} is discussed in detail in \Chapt~\ref{chapt:gettingstarted}. Note that all functions but {\tt glbChiSys} are using minimizers which have to be initialized with \GLB{SetInputErrors} and \GLB{SetStartingValues} first.}
\end{table}

@@

 glbSetStartingValues(true\_values); \\

@@

 chi2 = glbChiTheta(fit\_values,minimum,GLB\_ALL); \\ 

@@

 chi2 = glbChiThetaDelta(fit\_values,minimum,GLB\_ALL); \\

@@

  glbSetStartingValues(starting\_values);\\

@@

 chi2 = glbChiTheta(fit\_values,minimum,0);\\ 

@@

  chi2b = glbChiTheta(fit\_values,minimum,1); \\ 

@@

 chi2sum = glbChiTheta(fit\_values,minimum,GLB\_ALL); \\ 

@@

\index{norm}{Installation}
In this first chapter of the user's manual, we assume that the \GLOBES\ 
software is readily installed on your computer system. For the installation,
see \App~\ref{app:installation} and the {\tt INSTALL} file in the
software package. We demonstrate how to load pre-defined experiments 
and introduce the basic concepts of \GLOBES . We do not go
into details of the programming language, which means that standard parts
of the program code common to all of the examples in the following chapters 
are, in general, omitted.
An example of a minimal \GLOBES\ program in C can be found on 
page~\pageref{ex:c}. Furthermore, the files of the examples in this part 
can be found in the \verb+example+ subdirectory of your \GLOBES\ distribution.
 \index{norm}{Examples} 
After the installation of \GLOBES, they can be compiled using 
the \verb+Makefile+ in the \verb+examples+
 directory. The  \verb+Makefile+ has been correctly setup by the 
\verb+configure+
script to take into account details of the installation on your system. Thus
you've just to type \verb+make+ and you're done.
\footnote{The  data files (\AEDL\ and supporting files) needed by the
examples are already in place.} This \verb+Makefile+ very well serves
as a template for your own applications.
\index{norm}{Compilation!of application programs}
\index{norm}{Build process| \see{Compilation}}

@@

\example{Using \GLOBES\ with C}{\label{ex:c}
\index{norm}{C-Code}
\index{norm}{Program}
Here comes the C-code skeleton, which is
(more or less) common to all of our \GLOBES\ examples:
\begin{quote}
{\tt {\footnotesize
\#include <stdio.h> \\
\#include <stdlib.h> \\
\#include <math.h> \\
\#include <string.h> \\
\\
\#include <globes/globes.h> \hspace*{0.5cm} /* Include GLoBES library */ \\
\\
\#include "myio.h" \hspace*{0.5cm} /* Include "housemade" I/O-routines */ \\
\\
/* If filename given, write to file; if empty, to screen: */ \\
char MYFILE[]="testX.dat"; \\
\\
int main(int argc, char *argv[]) \\
\{  \\
\\
 \hspace*{0.5cm} glbInit(argv[0]); \hspace*{0.5cm} /* Initialize GLoBES library */ \\
\\  
  \hspace*{0.5cm} glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0], \\
  \hspace*{1cm} \&glb\_num\_of\_exps); \hspace*{0.5cm} /* Initialize experiment NuFact.glb */\\
\\  
  \hspace*{0.5cm}  /* Initialize housemade output function */\\
  \hspace*{0.5cm} 
   InitOutput(MYFILE,"Format: ... ... ... $\backslash$n"); \\
\\  
  \hspace*{0.5cm} /* Initialize parameter vector(s) */ \\
  \hspace*{0.5cm} glb\_params true\_values = glbAllocParams(); \\
  \hspace*{0.5cm} /* ... */ \\
\\  
 \hspace*{0.5cm} /* Assign: theta12,theta13,theta23,deltacp,dm2solar,dm2atm */ \\
   \hspace*{0.5cm}  
     glbDefineParams(true\_values,\\
     \hspace*{1.5cm}asin(sqrt(0.8))/2,asin(sqrt(0.001))/2,M\_PI/4,M\_PI/2,7e-5,2e-3); \\

@@

\underline{$\beta$-Beams:} & \multicolumn{3}{l}{In preparation} \\[0.1cm]
\hline
\end{tabular}
\end{center}
\mycaption{\label{tab:experiments} \index{norm}{Experiment files (table)}
\index{norm}{glb files@{\tt glb}-files} 
Pre-defined experiment prototypes, their filenames (to be used in \GLB{InitExperiment}), their short descriptions, and the references in which they are originally used and discussed (except from minor modifications, such as a
different implementation of the energy threshold function). Note that
some of these experiments are outdated in terms of integrated luminosities,
baseline, fluxes, efficiencies, or other factors. In any case these file are installed along with \GLOBES.}
\end{table}

@@

\index{norm}{Experiment!number of}
In principle, the \GLOBES\ user interface can currently handle up to 32 of different long-baseline experiments simultaneously, where the number
of existing experiment definition files can, of course, be unlimited. This means that their $\Delta \chi^2$-values are added {\em after} the minimization over the systematics parameters, and {\em before} any minimization over the oscillation parameters. Note that each experiment
assumes a specific matter density profile, which means
that it makes sense to simulate different operation modes within one
experiment definition, and physically different baselines in different
definitions. For details of the rate computation and
simulation techniques, we refer at this place to \Part~\ref{part:2}. Though
 the simplest case of simulating one experiment may be most often used, 
 using more than one experiments are useful in many cases. For example, combinations of experiments can be tested for
complementarity and competitiveness by equal means within one program.
In general, many \GLOBES\ functions take the experiment number as
a parameter, which runs from $0$ to \GLB{\_num\_of\_exps}-1 in the order of their initialization in the program.\footnote{Note that
the global variable {\tt glb\_num\_of\_exps} must not be modified by the
user.} In addition, using the parameter value \GLBC{GLB\_ALL} as
experiment number initiates a combined analysis of all loaded experiments.
\index{norm}{GLB ALL@{\tt GLB\_ALL}}
\index{constants}{{\tt GLB\_ALL}}

In general \GLBC{GLB\_ALL} can be used in many cases where there is an argument
selecting `$i$ out of $N$', \eg\ the 1st experiment out of 5, or the 5th
rule of 20. In those cases  using \GLBC{GLB\_ALL} is equivalent to
calling the corresponding function for \emph{all} $i$ in $N$  and `add' 
the effect of each
invocation, like in
\begin{quote}
\verb&for (i=0;i<N;i++) result += some_function(i);&\\
is the same as\\
\verb&result = some_function(GLB_ALL);&
\end{quote}
Here the meaning of `add' is, that whatever the desired result of calling
\verb+some_function+ is, this result is obtained for each $i$ in $N$, \eg\ 
setting the baseline in all experiments to a certain value or
or compute the $\chi^2$ for each experiment and return the total result.
There are however some functions where the action performed or the result
is so complex that is not possible or sensible to perform this for all $i$
in $N$. Calling these functions with  \GLBC{GLB\_ALL} as argument will 
in any case result in an exit status indicating
failure and the function will produce an 
error message\footnote{if the verbosity
level is set accordingly}.

\index{norm}{Experiment!list}
\index{norm}{Initialization!experiments}
For storing the experiments, \GLOBES\ uses the initially empty list of experiments \GLB{\_experiment\_list}. To add a pre-defined experiment to this list, one can use the function {\tt glbInitExperiment}:
\begin{function}
\index{norm}{Experiment initialization}
\GLBNS{InitExperiment}
{\tt int glbInitExperiment(char *inf, glb\_exp *in, int *counter)}
 adds a single experiment with the filename {\tt inf} to the list of currently loaded experiments. The {\tt counter} is a pointer to the 
 variable containing the number of experiments, and the experiment {\tt in}
 points to the beginning of the experiment list. The function returns
 zero if it was successful. 
\end{function}
Normally, a typical  call of {\tt glbInitExperiment} is 
\begin{quote}
{\tt glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0],\\  \hspace*{8cm} \&glb\_num\_of\_exps); }
\end{quote}
In this case, the experiment in the file {\tt NuFact.glb} is added to the internal global list of experiments, and the experiment counter is increased. The
experiment then has the number {\tt glb\_num\_of\_exps}-1. The elements
of the experiment list have the type \GLB{\_exp}, which the
user will not need to access directly in most cases. The experiment definition files, which usually end with {\tt .glb}, and any
supporting files, are first of all searched in the current directory, and then in the path given in the environment variable \verb+GLB_PATH+.
%
\index{norm}{GLB PATH@{\tt GLB\_PATH}}
\index{norm}{Environment variables!GLB PATH@{\tt GLB\_PATH}}
%

@@

Before we can set the simulated event rates or access any oscillation
parameters, we need to become familiar with the concept \GLOBES\ uses
for oscillation parameters. In order to transfer sets of oscillation
parameter vectors $(\theta_{12}, \theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm)$, the parameter type \GLB{\_params} is used. In general,
 this type is often
transferred to and from \GLOBES\ functions. Therefore, the memory 
for these vectors has to be reserved (allocated) before they can be used, 
and it has
to be returned (freed) afterwards. \GLOBES\ functions usually use the pointers
of the type {\tt glb\_params} for the input or output to the functions. As
an input parameter, the pointer has to be created and point towards a valid
parameter structure, where the oscillation parameters are read from. As
an output parameter, the pointer has to be created, too, and point towards
a structure which will contain the return values will be written to.
This parameter transfer concept seems to be very sophisticated, but, as we
will see in the next chapters, it hides a lot of complicated parameter 
mappings which otherwise need to be done by the user. For example, not
only the oscillation parameters are stored in the pointer structure, but also information on the matter
densities of all of the initialized experiments. Since \GLOBES\ treats
the matter density as a free parameter known with some external precision
to include matter density uncertainties,
the minimizers also use fit values and external errors for the matter
densities of all loaded experiments. More precisely, the matter density
profile of each experiment $i$ is multiplied by a scaling factor
$\hat{\rho_i}$, \index{norm}{Matter density!scaling factor}
\index{norm}{Matter density!profile} 
which is stored in the density information of {\tt glb\_params}. 
Each of
these scaling factors has $1.0$ as pre-defined value. Since it
is in most cases not necessary to change this value, the user does
not need to take care of it. For a constant matter density, it is simply the ratio of the matter density and the average matter density specified in the experiment definition, \ie , $\hat{\rho_i} \equiv \rho_i/\bar{\rho}_i$. For a matter density profile, it acts as an overall normalization factor: The matter density in each layer is multiplied by this factor. In most cases one wants to take a scaling factor of $1.0$ here, which simply means taking the matter density profile as it is given in the experiment definition. For the treatment of correlations, however,
an external precision of the scaling factor might be used to include the
correlations with the matter density uncertainty.
Note that the {\tt glb\_params}
structures must not be initialized before all experiments are loaded,
since the number of matter densities can only be determined after the
experiments are initialized. 
Similarly, any change in the number of experiments requires that the
parameter structures be re-initialized, \ie, freed and allocated again.

Another piece of information will be returned from the minimizers 
(\cf, \Chapt~\ref{chapt:correlations}) and transferred
into the {\tt glb\_params} structure is the number
of iterations used for the minimization, which is proportional to the
running time of the minimizer. In general, the user does not need to
access the elements in {\tt glb\_params} directly. A number of functions
is provided to handle these parameter structures:
\begin{function}
\GLBNS{AllocParams}
{\tt glb\_params glbAllocParams()} allocates the memory space
needed for a parameter vector and returns a pointer to it.
\end{function}
\begin{function}
\GLBNS{AllocParams}
{\tt void glbFreeParams(glb\_params stale)} frees the memory
needed for a parameter vector {\tt stale} and sets the pointer to {\tt NULL}.
\end{function}
\begin{function}
\GLBNS{DefineParams}
{\tt glb\_params glbDefineParams(glb\_params in, double theta12, double theta13,double theta23, double delta, double dms, double dma)} assigns
the complete set of oscillation parameters to the vector {\tt in}, which has
to be allocated before. The return value is the pointer to {\tt in} if 
the assignment was successful, and {\tt NULL} otherwise.
\end{function}
\begin{function}
 \GLBNS{CopyParams}
 {\tt glb\_params glbCopyParams(const glb\_params source, glb\_params dest)}
 copies the vector {\tt source} to the vector {\tt destination}. The return
 value is {\tt NULL} if the assignment was not successful.
\end{function}
\begin{function}
\GLBNS{PrintParams}
{\tt void glbPrintParams(FILE *stream, const glb\_params in)} prints the
parameters in {\tt in} to the file {\tt stream}. The oscillation
parameters, all density values, and the number of iterations are
printed as pretty output. Use {\tt stdout} for {\tt stream} if you want
to print to the screen.
\end{function}
In addition to these basic functions, there are functions to access
the individual parameters within the parameter vectors:
\begin{function}
\GLBNS{SetOscParams}
{\tt glb\_params glbSetOscParams(glb\_params in, double osc, int which)}
sets the oscillation parameter {\tt which} in the structure {\tt in}
to the value {\tt osc}. If the assignment was unsuccessful, the 
function returns {\tt NULL}.
\end{function}
\begin{function}
\GLBNS{GetOscParams}
{\tt double glbGetOscParams(glb\_params in, int which)}
returns the value of the oscillation parameter {\tt which} in the 
structure {\tt in}.
\end{function}
In both of these functions, the parameter {\tt which} runs from
$0$ to $5$, where the parameters in \GLOBES\ always have the order $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, $\deltacp$, $\sdm$, $\ldm$.
Alternatively to the number, the constants \GLBC{GLB\_THETA\_12}, \GLBC{GLB\_THETA\_13}, \GLBC{GLB\_THETA\_23}, \GLBC{GLB\_DELTA\_CP}, 
\GLBC{GLB\_DM\_SOL}, or \GLBC{GLB\_DM\_ATM} can be used.

@@

Other examples for projections onto a parameter axis while keeping exactly one parameter fixed are \GLB{ChiTheta23}, \GLB{ChiDm}, and \GLB{ChiDms}, which
can be found in \Tab~\ref{tab:stdfunctions} on page~\pageref{tab:stdfunctions}.

@@

The complete code can be found as {\tt example4.c} with the software,
which consists of many of the applications from the earlier examples.
In addition, it uses a little trick: It avoids falling into the wrong
 solution with {\tt glbChiTheta}
by using the fit value of $\deltacp$ from the step before as prediction
of the position of the current calculation.

>>>>

TOUR ITSELF (+EXAMPLE FILE) TO BE UPDATED. JK: PLEASE LET ME KNOW WHEN {\tt example-tour.c} IS FIXED. MAYBE WE SHOULD  ALSO UPDATE THE PARAMTER VALUES [WW].

@@

\begin{table}[tpb]
\begin{center}
\begin{tabular}{p{3.8cm}p{3.8cm}p{7cm}}
\hline
Function & Purpose & Parameters \ra\ Result \\
\hline
\multicolumn{3}{l}{{\bf Systematics only:}} \\
\GLB{ChiSys} & $\chi^2$ with systematics only  & ({\tt glb\_params in, int exp, int rule}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projections onto axes:}} \\
\GLB{ChiTheta13} & Projection onto $\theta_{13}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiDelta} & Projection onto $\deltacp$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiTheta23} & Projection onto $\theta_{23}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiDm31} & Projection onto $\ldm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiDm21} & Projection onto $\sdm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projection onto plane:}} \\
\GLB{ChiTheta13Delta} & Projection onto $\theta_{13}$-$\deltacp$-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projection onto any hyper-plane:}} \\
\GLB{ChiNP} & Projection onto any $n$-dimensional hyper-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \newline
Needs \GLB{SetProjection} before! \\[0.2cm]
\multicolumn{3}{l}{{\bf Localization of degeneracies:}} \\
\GLB{ChiAll} & (Local) Minimization over all parameters  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\
\hline
\end{tabular}
\end{center}
\caption{\label{tab:stdfunctions} \index{norm}{Standard functions (table)} The \GLOBES\ standard function to obtain a $\chi^2$-value with systematics only or systematics and correlations. The parameters {\tt rule} and {\tt exp}
can either be \GLBC{GLB\_ALL} for all initialized experiment or the
experiment number ($0$ to \GLB{\_num\_of\_exps}-1) for a specific experiment. The format of \GLB{\_params} is discussed in detail in \Chapt~\ref{chapt:gettingstarted}. Note that all functions but {\tt glbChiSys} are using minimizers which have to be initialized with \GLB{SetInputErrors} and \GLB{SetCentralValues} first.}
\end{table}

@@

 glbSetCentralValues(true\_values); \\

@@

 chi2 = glbChiTheta13(fit\_values,minimum,GLB\_ALL); \\ 

@@

 chi2 = glbChiTheta13Delta(fit\_values,minimum,GLB\_ALL); \\

@@

  glbSetCentralValues(starting\_values);\\

@@

 chi2 = glbChiTheta13(fit\_values,minimum,0);\\ 

@@

  chi2b = glbChiTheta13(fit\_values,minimum,1); \\ 

@@

 chi2sum = glbChiTheta13(fit\_values,minimum,GLB\_ALL); \\ 

@@

\index{norm}{Installation}
In this first chapter of the user's manual, we assume that the \GLOBES\ 
software is readily installed on your computer system. For the installation,
see \App~\ref{app:installation} and the {\tt INSTALL} file in the
software package. We demonstrate how to load pre-defined experiments 
and introduce the basic concepts of \GLOBES . We do not go
into details of the programming language, which means that standard parts
of the program code common to all of the examples in the following chapters 
are, in general, omitted.
An example of a minimal \GLOBES\ program in C can be found on 
page~\pageref{ex:c}. Furthermore, the files of the examples in this part 
can be found in the \verb+example+ subdirectory of your \GLOBES\ distribution.
 \index{norm}{Examples} 
After the installation of \GLOBES, they can be compiled using 
the \verb+Makefile+ in the \verb+examples+
 directory. The  \verb+Makefile+ has been correctly setup by the 
\verb+configure+
script to take into account details of the installation on your system. Thus
you've just to type \verb+make+ and you're done.\footnote{The  
data files (\AEDL\ and supporting files) needed by the
examples are already in place.} This \verb+Makefile+ very well serves
as a template for your own applications.
\index{norm}{Compilation!of application programs}
\index{norm}{Build process| \see{Compilation}}

@@

\example{Using \GLOBES\ with C}{\label{ex:c}
\index{norm}{C-Code}
\index{norm}{Program}
Here comes the C-code skeleton, which is
(more or less) common to all of our \GLOBES\ examples:
\begin{quote}
{\tt {\footnotesize
\#include <stdio.h> \\
\#include <stdlib.h> \\
\#include <math.h> \\
\#include <string.h> \\
\\
\#include <globes/globes.h> \hspace*{0.5cm} /* Include GLoBES library */ \\
\\
\#include "myio.h" \hspace*{0.5cm} /* Include "housemade" I/O-routines */ \\
\\
/* If filename given, write to file; if empty, to screen: */ \\
char MYFILE[]="testX.dat"; \\
\\
int main(int argc, char *argv[]) \\
\{  \\
\\
 \hspace*{0.5cm} glbInit(argv[0]); \hspace*{0.5cm} /* Initialize GLoBES library */ \\
\\  
  \hspace*{0.5cm} glbInitExperiment("NFstandard.glb",\&glb\_experiment\_list[0], \\
  \hspace*{1cm} \&glb\_num\_of\_exps); \hspace*{0.5cm} /* Initialize experiment NFstandard.glb */\\
\\  
  \hspace*{0.5cm}  /* Initialize housemade output function */\\
  \hspace*{0.5cm} 
   InitOutput(MYFILE,"Format: ... ... ... $\backslash$n"); \\
\\  
  \hspace*{0.5cm} /* Initialize parameter vector(s) */ \\
  \hspace*{0.5cm} glb\_params true\_values = glbAllocParams(); \\
  \hspace*{0.5cm} /* ... */ \\
\\  
 \hspace*{0.5cm} /* Assign: theta12,theta13,theta23,deltacp,dm2solar,dm2atm */ \\
   \hspace*{0.5cm}  
     glbDefineParams(true\_values,\\
     \hspace*{1.5cm}asin(sqrt(0.8))/2,asin(sqrt(0.001))/2,M\_PI/4,M\_PI/2,7e-5,2e-3); \\

@@

\underline{$\beta$-Beams:} & \multicolumn{3}{l}{In preparation} \\[0.1cm]
\hline
\end{tabular}
\end{center}
\mycaption{\label{tab:experiments} \index{norm}{Experiment files (table)}
\index{norm}{glb files@{\tt glb}-files} 
TO BE UPDATED BY MR; POSSIBLY MORE DETAILED DESCRIPTION IN APPENDIX [WW]. Pre-defined experiment prototypes, their filenames (to be used in \GLB{InitExperiment}), their short descriptions, and the references in which they are originally used and discussed (except from minor modifications, such as a
different implementation of the energy threshold function). Note that
some of these experiments are outdated in terms of integrated luminosities,
baseline, fluxes, efficiencies, or other factors. In any case these file are installed along with \GLOBES.}
\end{table}

@@

\index{norm}{Experiment!number of}
In principle, the \GLOBES\ user interface can currently handle up to 32 of different long-baseline experiments simultaneously, where the number
of existing experiment definition files can, of course, be unlimited. This means that their $\Delta \chi^2$-values are added {\em after} the minimization over the systematics parameters, and {\em before} any minimization over the oscillation parameters. Note that each experiment
assumes a specific matter density profile, which means
that it makes sense to simulate different operation modes within one
experiment definition, and physically different baselines in different
definitions. For details of the rate computation and
simulation techniques, we refer at this place to \Part~\ref{part:2}. Though
 the simplest case of simulating one experiment may be most often used, 
 using more than one experiments are useful in many cases. For example, combinations of experiments can be tested for
complementarity and competitiveness by equal means within one program.
In general, many \GLOBES\ functions take the experiment number as
a parameter, which runs from $0$ to \GLB{\_num\_of\_exps}-1 in the order of their initialization in the program.\footnote{Note that
the global variable {\tt glb\_num\_of\_exps} must not be modified by the
user.} In addition, using the parameter value \GLBC{GLB\_ALL} as
experiment number (for example, in the {\tt glbChi...} functions) initiates a combined analysis of all loaded experiments.
\index{norm}{GLB ALL@{\tt GLB\_ALL}}
\index{constants}{{\tt GLB\_ALL}}

% THIS PART IS CONFUSING, BECAUSE THE EXPLAINED BEHAVIOR IS DIFFERENT FROM THE glbChi-functions
% (THE MARGINALIZATION IS PERFORMED AFTER THE CHI2 IS ADDED! ONE MAY GET A DIFFERENT IMPRESSION
% FROM THIS DESCRIPTION). THEREFORE, I HAVE REMOVED IT [WW].
%
% In general \GLBC{GLB\_ALL} can be used in many cases where there is an argument
% selecting `$i$ out of $N$', \eg\ the 1st experiment out of 5, or the 5th
% rule of 20. In those cases  using \GLBC{GLB\_ALL} is equivalent to
% calling the corresponding function for \emph{all} $i$ in $N$  and `add' 
% the effect of each
% invocation, such as in
% \begin{quote}
% \verb&for (i=0;i<N;i++) result += some_function(i);&\\
% is the same as\\
% \verb&result = some_function(GLB_ALL);&
% \end{quote}
% Here the meaning of `add' is, that whatever the desired result of calling
% \verb+some_function+ is, this result is obtained for each $i$ in $N$, \eg\ 
% setting the baseline in all experiments to a certain value or
% or compute the $\chi^2$ for each experiment and return the total result.
% There are however some functions where the action performed or the result
% is so complex that is not possible or sensible to perform this for all $i$
% in $N$. Calling these functions with  \GLBC{GLB\_ALL} as argument will 
% in any case result in an exit status indicating
% failure and the function will produce an 
% error message\footnote{if the verbosity
% level is set accordingly}.

\index{norm}{Experiment!list}
\index{norm}{Initialization!experiments}
For storing the experiments, \GLOBES\ uses the initially empty list of experiments \GLB{\_experiment\_list}. To add a pre-defined experiment to this list, one can use the function {\tt glbInitExperiment}:
\begin{function}
\index{norm}{Experiment initialization}
\GLBNS{InitExperiment}
{\tt int glbInitExperiment(char *infile, glb\_exp *ptr, int *counter)}
 adds a single experiment with the filename {\tt infile} to the list of currently loaded experiments. The {\tt counter} is a pointer to the 
 variable containing the number of experiments, and the experiment {\tt ptr}
 points to the beginning of the experiment list. The function returns
 zero if it was successful. 
\end{function}
Normally, a typical  call of {\tt glbInitExperiment} is 
\begin{quote}
{\tt glbInitExperiment("NFstandard.glb",\&glb\_experiment\_list[0],\\  \hspace*{8cm} \&glb\_num\_of\_exps); }
\end{quote}
In this case, the experiment in the file {\tt NFstandard.glb} is added to the internal global list of experiments, and the experiment counter is increased. The
experiment then has the number {\tt glb\_num\_of\_exps}-1. The elements
of the experiment list have the type \GLB{\_exp}, which the
user will not need to access directly in most cases. The experiment definition files, which usually end with {\tt .glb}, and any
supporting files, are first of all searched in the current directory, and then in the path given in the environment variable \verb+GLB_PATH+.
%
\index{norm}{GLB PATH@{\tt GLB\_PATH}}
\index{norm}{Environment variables!GLB PATH@{\tt GLB\_PATH}}
%

@@

Before we can set the simulated event rates or access any oscillation
parameters, we need to become familiar with the concept \GLOBES\ uses
for oscillation parameters. In order to transfer sets of oscillation
parameter vectors $(\theta_{12}, \theta_{13}, \theta_{23}, \deltacp, \sdm, \ldm)$
as well as some other information, the parameter type \GLB{\_params} is used. In general,
 this type is often
transferred to and from \GLOBES\ functions. Therefore, the memory 
for these vectors has to be reserved (allocated) before they can be used, 
and it has
to be returned (freed) afterwards. \GLOBES\ functions usually use the pointers
of the type {\tt glb\_params} for the input or output to the functions. As
an input parameter, the pointer has to be created and point towards a valid
parameter structure, where the oscillation parameters are read from. As
an output parameter, the pointer has to be created, too, and point towards
a structure which the return values will be written to.
This parameter transfer concept seems to be very sophisticated, but, as we
will see in the next chapters, it hides a lot of complicated parameter 
mappings which otherwise need to be done by the user. For example, not
only the oscillation parameters are stored in the pointer structure, but also information on the matter
densities of all of the initialized experiments. Since \GLOBES\ treats
the matter density as a free parameter known with some external precision
to include matter density uncertainties,
the minimizers also use fit values and external errors for the matter
densities of all loaded experiments. More precisely, the matter density
profile of each experiment $i$ is multiplied by a scaling factor
$\hat{\rho_i}$, \index{norm}{Matter density!scaling factor}
\index{norm}{Matter density!profile} 
which is stored in the density information of {\tt glb\_params}. 
Each of
these scaling factors has $1.0$ as pre-defined value. Since it
is in most cases not necessary to change this value, the user does
not need to take care of it. For a constant matter density, it is simply the ratio of the matter density and the average matter density specified in the experiment definition, \ie , $\hat{\rho_i} \equiv \rho_i/\bar{\rho}_i$. For a matter density profile, it acts as an overall normalization factor: The matter density in each layer is multiplied by this factor. In most cases one wants to take a scaling factor of $1.0$ here, which simply means taking the matter density profile as it is given in the experiment definition. For the treatment of correlations, however,
an external precision of the scaling factor might be used to include the
correlations with the matter density uncertainty.
Note that the {\tt glb\_params}
structures must not be initialized before all experiments are loaded,
since the number of matter densities can only be determined after the
experiments are initialized. 
Similarly, any change in the number of experiments requires that the
parameter structures be re-initialized, \ie, freed and allocated again.

Another piece of information will be returned from the minimizers 
(\cf, \Chapt~\ref{chapt:correlations}) and transferred
into the {\tt glb\_params} structure is the number
of iterations used for the minimization, which is proportional to the
running time of the minimizer. In general, the user does not need to
access the elements in {\tt glb\_params} directly. A number of functions
is provided to handle these parameter structures:
\begin{function}
\GLBNS{AllocParams}
{\tt glb\_params glbAllocParams()} allocates the memory space
needed for a parameter vector and returns a pointer to it.
\end{function}
\begin{function}
\GLBNS{AllocParams}
{\tt void glbFreeParams(glb\_params stale)} frees the memory
needed for a parameter vector {\tt stale} and sets the pointer to {\tt NULL}.
\end{function}
\begin{function}
\GLBNS{DefineParams}
{\tt glb\_params glbDefineParams(glb\_params in, double theta12, double theta13,double theta23, double delta, double dm21, double dm31)} assigns
the complete set of oscillation parameters to the vector {\tt in}, which has
to be allocated before. The return value is the pointer to {\tt in} if 
the assignment was successful, and {\tt NULL} otherwise.
\end{function}
\begin{function}
 \GLBNS{CopyParams}
 {\tt glb\_params glbCopyParams(const glb\_params source, glb\_params dest)}
 copies the vector {\tt source} to the vector {\tt destination}. The return
 value is {\tt NULL} if the assignment was not successful.
\end{function}
\begin{function}
\GLBNS{PrintParams}
{\tt void glbPrintParams(FILE *stream, const glb\_params in)} prints the
parameters in {\tt in} to the file {\tt stream}. The oscillation
parameters, all density values, and the number of iterations are
printed as pretty output. Use {\tt stdout} for {\tt stream} if you want
to print to the screen.
\end{function}
In addition to these basic functions, there are functions to access
the individual parameters within the parameter vectors:
\begin{function}
\GLBNS{SetOscParams}
{\tt glb\_params glbSetOscParams(glb\_params in, double osc, int which)}
sets the oscillation parameter {\tt which} in the structure {\tt in}
to the value {\tt osc}. If the assignment was unsuccessful, the 
function returns {\tt NULL}.
\end{function}
\begin{function}
\GLBNS{GetOscParams}
{\tt double glbGetOscParams(glb\_params in, int which)}
returns the value of the oscillation parameter {\tt which} in the 
structure {\tt in}.
\end{function}
In both of these functions, the parameter {\tt which} runs from
$0$ to $5$ (or the number of oscillation parameters-1), where the parameters in \GLOBES\ always have the order $\theta_{12}$, $\theta_{13}$, $\theta_{23}$, $\deltacp$, $\sdm$, $\ldm$.
Alternatively to the number, the constants \GLBC{GLB\_THETA\_12}, \GLBC{GLB\_THETA\_13}, \GLBC{GLB\_THETA\_23}, \GLBC{GLB\_DELTA\_CP}, 
\GLBC{GLB\_DM\_SOL}, or \GLBC{GLB\_DM\_ATM} can be used.

@@

Other examples for projections onto a parameter axis while keeping exactly one parameter fixed are \GLB{ChiTheta23}, \GLB{ChiDm31}, and \GLB{ChiDm21}, which
can be found in \Tab~\ref{tab:stdfunctions} on page~\pageref{tab:stdfunctions}.

@@

The complete code can be found as {\tt example4.c} with the software,
which consists of many of the applications from the earlier examples.
In addition, it uses a little trick: It avoids falling into the wrong
 solution with {\tt glbChiTheta13}
by using the fit value of $\deltacp$ from the step before as prediction
of the position of the current calculation.
